\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2016-12-05 Response.tex

\begin{correction}
    We choose the domain \( D \) to exclude the undesired critical points of \( A(R) \) as well as the infeasible regions defined by the state inequality constraints. 
    Furthermore, we can show that the set \( D \) is an open neighborhood of the desired attitude configuration, \( R_d\).
\end{correction}
\item \textit{Typo in eq. (40).}

The equation has been modified as:
\begin{correction}
\begin{gather}
    0 < c < \frac{4 k_R k_\Omega}{k_\Omega^2 + 4 k_R \lambda_M H} ,
\end{gather}
\end{correction}

\item \textit{Abuse of notation 'hat' and 'vee' in equations (6)-(8)}

The \( \wedge : \R^3 \to \so \) and \( \vee : \so \to \R^3 \) operators are used according to the convention that is common in the literature~\cite{lee2010,lee2011a}.
In this case, the \textit{hat} operator is equivalently represented as \( \hat{x} = \parenth{x}^\wedge \).
This representation enables a consistent notation for both short/simple terms, such as \( \hat{x} \) and more complicated terms such as \(\parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee\). 
Finally, this dual usage of the notation drastically improves the quality of the resulting typography.

### 2016-12-05 manuscript.tex
Try to get author bios onto the last page
Move figure in paper

\documentclass[fleqn]{IJCAS}  % Comment this line out if you need a4paper
\usepackage{ijcas_packages}

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\end{multicols}
\begin{figure}
    \centering 
    \subfigure[Attitude error vector \(e_R\) components \label{fig:eR_con}]{\includegraphics[width=0.3\columnwidth]{eR_con}  }~
    \subfigure[Configuration error \( \Psi \) \label{fig:Psi_con}]{\includegraphics[width=0.3\columnwidth]{figures/Psi_con.pdf} }
    \subfigure[Angle to each constraint \label{fig:con_angles_con}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angles_con.pdf} } 
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
\begin{multicols}{2}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\end{multicols}
\begin{figure} 
	\centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.4\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to each constraint \label{fig:con_angles}]{\includegraphics[width=0.4\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\includegraphics[width=0.4\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
	\caption{Attitude stabilization with adaptive update law}
	\label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
Simulation results without the adaptive update law are shown in~\cref{fig:con}.
\Cref{fig:eR_con} shows each component of the attitude error vector,~\cref{eqn:eR}, over the simulation time span.
\Cref{fig:Psi_con} shows the  magnitude of the combined error function,~\cref{eqn:psi}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
In spite of the uncompensated disturbance, the system is able to avoid the constrained regions as shown in~\cref{fig:con_angles_con}.
The angle to each of the constraints, which is measured in degrees and given by \( \arccos(r^T R^T v_i) \), is always greater than the specified angle, \( \theta_i \), in~\Cref{tab:constraints}.

@@

% singularities and the problems involved with using Euler angles
Euler angle sequences are a minimum, three-parameter set of angles which describe the transformation between two reference frames.
Euler angles represent any general rotation as a sequence of three intermediate rotations.
By convention, there are \num{24} possible Euler angle sequences for any given rotation.
In addition, Euler angles are a minimum representation, as only three angles, and the associated sequence, are required to describe the three angular degrees of freedom of the rigid body.
However, there is great ambiguity in the representation of the attitude as there are many equivalent Euler angle sequences for a given attitude of the system.
Therefore, great care must be taken in the control system design to ensure that a consistent sequence is used. 
Furthermore, it has been shown that no minimal attitude representation can describe orientations both globally and without singularities~\cite{hughes2004,bhat2000}.
These singularities can cause significant difficulties during control design and hardware implementation.

To demonstrate the effect of the kinematic singularities inherent with Euler angles we will represent the attitude of the body fixed reference frame, \( \vecbf{b}_i \), with respect to the inertial frame, \( \vecbf{e}_i\), in terms of the \( 3-1-3\) Euler angle sequence.
More explicitly, this corresponds to the rotation sequence \( \theta_1 \vecbf{b}_3 , \theta_2 \vecbf{b}_1, \theta_3 \vecbf{b}_3 \).
The rotation matrix, \( R(\theta_1, \theta_2, \theta_3) \), corresponding to this sequence is 
\begin{align}\label{eq:euler313}
    \begin{bmatrix}
        -s_1 c_2 s_3 + c_3 c_1 & -s_1 c_2 c_3 - s_3 c_1 & s_1s_2 \\
        c_1 c_2 s_3 + c_3 s_1 & c_1 c_2 c_3 - s_3 s_1 & - c_1 s_2 \\
        s_2 s_3 & s_2 c_3 & c_2
    \end{bmatrix} ,
\end{align}
where \( s_i, c_i \) represent \( \sin \theta_i, \cos \theta_i \) for \( i = \braces{1,2,3}\).
Using this representation that kinematic differential equations for the associated Euler angles are given as
\begin{align}\label{eq:euler313_diff}
    \begin{bmatrix}
        \dot{\theta}_1 \\ \dot{\theta}_2 \\ \dot{\theta}_3 
    \end{bmatrix}
    =
    \begin{bmatrix}
        \slfrac{\parenth{\Omega_1 s_3 + \Omega_2 c_3}}{s_2} \\
        \Omega_1 c_3 - \Omega_2 s_3 \\
        -\slfrac{\parenth{\Omega_1 s_3 + \omega_2 c_3}c_2}{s_2} + \Omega_3
    \end{bmatrix} .
\end{align}
From~\cref{eq:euler313_diff}, it is immediately clear that a singularity exists when \( \sin \theta_2 = 0 \) or \( \theta_2 = 0, \pm \pi \). 
In the vicinity of the singularity, the angular velocities of the Euler angles will tend to approach \( \pm \infty \) and the angular velocities will experience instantaneous sign changes.
Furthermore, all Euler angle sequences will exhibit a similar singularity at either \( \theta_2 = 0, \pm \pi \) or \( \theta_2 = \pm \frac{\pi}{2}, \pm \frac{3\pi}{2} \) and switching the sequence does not alleviate the issue.
As a result, Euler angles are not appropriate for systems which experience large angular rotations, such as those demonstrated in~\cref{fig:adapt}, or control systems which rely on the angular velocities \( \theta_i \).

@@

%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

@@

\biography{profile}{Shankar Kulumani}
    {is a PhD Candidate in the Department of Mechanical and Aerospace Engineering at George Washington University. 
    He received his bachelor's degree in Astronautical Engineering from the United States Air Force Academy in 2009 and his master's in Aeronautical and Astronautical Engineering from Purdue University in 2013.
    % He previously served as Deputy Space Vehicles Lead for the ORS-1 (Operationally Responsive Space ) satellite program, and as Lead Test Engineer in the Guidance, Navigation, and Control group at Air Force Research Laboratory located at Kirtland AFB, NM.
    % He is now currently a reserve air force officer where he serves as Threat Systems Engineer for the Missle and Space Intelligence Center, Defense Intelligence Agency.
    His current research interests include the application of geometric mechanics and control to aerospace systems. 
    }

\biography{Lee}{Taeyoung Lee}
    {is an associate professor of Department of Mechanical and Aerospace Engineering at the George Washington University. 
    He received his doctoral degree in Aerospace Engineering and his master's degree in Mathematics at the University of Michigan in 2008. 
    His research interests include computational geometric mechanics and control of complex systems.}    
\clearafterbiography\relax
\clearafterbiography\relax
\end{multicols}

>>>>

\documentclass[10pt,fleqn]{IJCAS}  % Comment this line out if you need a4paper
\usepackage{ijcas_packages}

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\end{multicols}
\begin{figure}
    \centering 
    \subfigure[Attitude error vector \(e_R\) components \label{fig:eR_con}]{\includegraphics[width=0.3\columnwidth]{eR_con}  }~
    \subfigure[Configuration error \( \Psi \) \label{fig:Psi_con}]{\includegraphics[width=0.3\columnwidth]{figures/Psi_con.pdf} }
    \subfigure[Angle to each constraint \label{fig:con_angles_con}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angles_con.pdf} } 
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
\begin{multicols}{2}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
\Cref{fig:eR_con} shows each component of the attitude error vector,~\cref{eqn:eR}, over the simulation time span.
\Cref{fig:Psi_con} shows the  magnitude of the combined error function,~\cref{eqn:psi}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
In spite of the uncompensated disturbance, the system is able to avoid the constrained regions as shown in~\cref{fig:con_angles_con}.
The angle to each of the constraints, which is measured in degrees and given by \( \arccos(r^T R^T v_i) \), is always greater than the specified angle, \( \theta_i \), in~\Cref{tab:constraints}.

@@

% singularities and the problems involved with using Euler angles
Euler angle sequences are a minimum, three-parameter set of angles which describe the transformation between two reference frames.
Euler angles represent any general rotation as a sequence of three intermediate rotations.
By convention, there are \num{24} possible Euler angle sequences for any given rotation.
In addition, Euler angles are a minimum representation, as only three angles, and the associated sequence, are required to describe the three angular degrees of freedom of the rigid body.
However, there is great ambiguity in the representation of the attitude as there are many equivalent Euler angle sequences for a given attitude of the system.
Therefore, great care must be taken in the control system design to ensure that a consistent sequence is used. 
Furthermore, it has been shown that no minimal attitude representation can describe orientations both globally and without singularities~\cite{hughes2004,bhat2000}.
These singularities can cause significant difficulties during control design and hardware implementation.
\end{multicols}
\begin{figure} 
  \centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.4\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to each constraint \label{fig:con_angles}]{\includegraphics[width=0.4\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\includegraphics[width=0.4\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
  \caption{Attitude stabilization with adaptive update law}
  \label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
To demonstrate the effect of the kinematic singularities inherent with Euler angles we will represent the attitude of the body fixed reference frame, \( \vecbf{b}_i \), with respect to the inertial frame, \( \vecbf{e}_i\), in terms of the \( 3-1-3\) Euler angle sequence.
More explicitly, this corresponds to the rotation sequence \( \theta_1 \vecbf{b}_3 , \theta_2 \vecbf{b}_1, \theta_3 \vecbf{b}_3 \).
The rotation matrix, \( R(\theta_1, \theta_2, \theta_3) \), corresponding to this sequence is 
\begin{align}\label{eq:euler313}
    \begin{bmatrix}
        -s_1 c_2 s_3 + c_3 c_1 & -s_1 c_2 c_3 - s_3 c_1 & s_1s_2 \\
        c_1 c_2 s_3 + c_3 s_1 & c_1 c_2 c_3 - s_3 s_1 & - c_1 s_2 \\
        s_2 s_3 & s_2 c_3 & c_2
    \end{bmatrix} ,
\end{align}
where \( s_i, c_i \) represent \( \sin \theta_i, \cos \theta_i \) for \( i = \braces{1,2,3}\).
Using this representation that kinematic differential equations for the associated Euler angles are given as
\begin{align}\label{eq:euler313_diff}
    \begin{bmatrix}
        \dot{\theta}_1 \\ \dot{\theta}_2 \\ \dot{\theta}_3 
    \end{bmatrix}
    =
    \begin{bmatrix}
        \slfrac{\parenth{\Omega_1 s_3 + \Omega_2 c_3}}{s_2} \\
        \Omega_1 c_3 - \Omega_2 s_3 \\
        -\slfrac{\parenth{\Omega_1 s_3 + \omega_2 c_3}c_2}{s_2} + \Omega_3
    \end{bmatrix} .
\end{align}
From~\cref{eq:euler313_diff}, it is immediately clear that a singularity exists when \( \sin \theta_2 = 0 \) or \( \theta_2 = 0, \pm \pi \). 
In the vicinity of the singularity, the angular velocities of the Euler angles will tend to approach \( \pm \infty \) and the angular velocities will experience instantaneous sign changes.
Furthermore, all Euler angle sequences will exhibit a similar singularity at either \( \theta_2 = 0, \pm \pi \) or \( \theta_2 = \pm \frac{\pi}{2}, \pm \frac{3\pi}{2} \) and switching the sequence does not alleviate the issue.
As a result, Euler angles are not appropriate for systems which experience large angular rotations, such as those demonstrated in~\cref{fig:adapt}, or control systems which rely on the angular velocities \( \theta_i \).

@@

@@

\biography{profile}{Shankar Kulumani}
    {is a PhD Candidate in the Department of Mechanical and Aerospace Engineering at George Washington University. 
    He received his bachelor's degree in Astronautical Engineering from the United States Air Force Academy in 2009 and his master's in Aeronautical and Astronautical Engineering from Purdue University in 2013.
    % He previously served as Deputy Space Vehicles Lead for the ORS-1 (Operationally Responsive Space ) satellite program, and as Lead Test Engineer in the Guidance, Navigation, and Control group at Air Force Research Laboratory located at Kirtland AFB, NM.
    % He is now currently a reserve air force officer where he serves as Threat Systems Engineer for the Missle and Space Intelligence Center, Defense Intelligence Agency.
    His current research interests include the application of geometric mechanics and control to aerospace systems. 
    }
\biography{Lee}{Taeyoung Lee}
    {is an associate professor of Department of Mechanical and Aerospace Engineering at the George Washington University. 
    He received his doctoral degree in Aerospace Engineering and his master's degree in Mathematics at the University of Michigan in 2008. 
    His research interests include computational geometric mechanics and control of complex systems.}    
\clearafterbiography\relax

\end{multicols}
